Dynamical modeling of tidal streams

### 2014-01-07 tex/ms.tex
streams: description of new figure

>>>>

\begin{figure}[t!]
  \includegraphics[width=0.48\textwidth,clip=]{gd1-hisigv_evol_xz.eps}
  \caption{Same as \figurename~\ref{fig:gd1_xz}, but for a simulation
    of a progenitor with a mass $M^p$ that is $1,000$ times larger and
    therefore a velocity dispersion $\sigma_v^p$ that is ten times
    larger. The model used to predict the stream track (solid line) is
    the same as that used in \figurename~\ref{fig:gd1_xz}, except that
    the model velocity dispersion \sigv\ has been increased by a
    factor of ten. The stream spreads over a much larger volume for a
    heavier progenitor, but the fiducial model still accurately
    predicts its location in phase space.}\label{fig:gd1-hisigv_xz}
%python plot_stream_hisigv.py ../tex/gd1-hisigv_evol_xz.ps
\end{figure}

\subsection{At what progenitor mass does the framework break down?}

I have illustrated the new framework for modeling tidal streams using
a cold, narrow stream with a progenitor velocity dispersion of only
$\sigma_v^p = 1.4\kms$ (and a progenitor mass of $M^p = 2 \times
10^4\msun$). For a tidal stream that is as cold as that, the offset
between the progenitor's orbit and the stream track as well as the
dispersion within the stream are small (we expect these relative
offsets to be approximately $\sigv/V_c$). Therefore, we can linearize
the $(\vecx,\vecv) \leftrightarrow (\veco,\veca)$ transformation over
the relevant range of $(\vecx,\vecv)$.  The stream dispersion and the
stream--progenitor-orbit offset scale approximately as mass$^{1/3}$,
such that for streams arising from the tidal disruption of small dwarf
galaxies, we expect the stream--progenitor-orbit offset to be around
$10\,\%$. In the fiducial model and for the simulated stream used in
this paper, the internal frequency dispersion in the stream is six
times smaller than the stream--progenitor-orbit offset. Therefore,
bridging the latter with a linearized frequency--angle transformation
is the more stringent constraint.

To test whether the framework presented applies to streams arising
from the tidal disruption of small dwarf galaxies, I have run the same
simulation as that described in \sectionname~\ref{sec:sim}, but for a
progenitor mass of $M^p=2\times 10^7\msun$ and a velocity dispersion
of $\sigma_v^p=14\kms$. The tidal radius of this progenitor is
$0.69\kpc$ and I use a softening parameter of $14\pc$ in this
case. The final snapshot of this simulation in Galactocentric $X$ and
$Z$ coordinates is shown in \figurename~\ref{fig:gd1-hisigv_xz}. I
predict the stream track for this simulated stream by taking the same
fiducial model as decribed in \sectionname~\ref{sec:fidmodel}, except
that the model velocity dispersion \sigv\ is increased by a factor of
ten. This track is shown in \figurename~\ref{fig:gd1-hisigv_xz}.

It is clear from \figurename~\ref{fig:gd1-hisigv_xz} that the fiducial
model as well as the action--angle approximations used in this paper
still work for heavier progenitors with masses $\approx
10^7\msun$. Similar experiments with progenitors with masses $\approx
10^8\msun$ show that the framework presented here can model the tidal
streams produced by such heavy progenitor's as well, although the
stream--progenitor-orbit offset becomes so large that the algorithm
for calculating the stream track in $(\vecx,\vecv)$ described in
\sectionname~\ref{sec:trackxv} may need to be improved. When the
stream--progenitor-orbit becomes so large that the assumption of a
linear $(\vecx,\vecv) \leftrightarrow (\veco,\veca)$ transformation
becomes invalid, we can improve the calculation of the stream track by
iterating the procedure described in \sectionname~\ref{sec:trackxv}
and illustrated in \figurename~\ref{fig:track}. After calculating the
stream track, we can calculate the $(\vecx,\vecv) \rightarrow
(\veco,\veca)$ transformation at the estimated track points and use
this mapping to refine the estimated trackpoints. A single iteration
of this procedure will likely suffice. With this procedure, streams
originating from progenitors up to masses of $10^9\msun$ should be
able to be modeled with the framework presented in this paper.

Above $M^p = 10^9\msun$ the effects of dynamical friction, the dwarf
galaxy's gravity, and non-linearity in the $\vecj \rightarrow \veco$
transformation are expected to start affecting the structure and
location of the stream \citep[\eg,][]{Sanders13a}. Therefore, streams
originating from the heaviest dwarf satellites will have to be modeled
with more intricate means.
